ms_results

### 2023-02-04 ms_figures.Rmd
Update analyses to exclude non-human food

```{r load_data}
# Set directories
datadir <- "/Volumes/jgephart/ARTIS/Outputs/ms_seafood_globalization/20220111"
outdir <- "/Volumes/jgephart/ARTIS/Outputs/ms_seafood_globalization/outputs_20220111"

@@

```{r load_drive_data}

pop <- read.csv("/Volumes/jgephart/ARTIS/Outputs/clean_metadata/fao_annual_pop.csv") %>% 
  left_join(country_metadata, by = c("iso3c")) %>%
  select(iso3c, year, "region" = "owid_region", pop)

@@

fig1b <- ggplot() +
  geom_line(data = percent_export_by_source, aes(x = year, y = percent_export, color = habitat_method)) +
  scale_color_manual(values = habitat_source_colors) +
  labs(y = "% Production Exported", x = "", color = "Source") +
  theme_bw()

@@

```{r fig2a}
# Imports by Region
fig2a <- artis_region %>% 
  ggplot(aes(x = year, y = live_weight_t/1000000, fill = exporter_region)) +
  geom_area() +
  labs(x = "", y = "", fill = "Exporter") +
  scale_fill_manual(values = region_colors) +
  facet_wrap(~importer_region, nrow = 1) +
  theme_bw() +
  theme(legend.position = "none",
        strip.background = element_blank(),
        strip.text.x = element_blank(),
        axis.text.x = element_blank(),
        plot.margin=unit(c(0.5,0,0,0), "cm"))

@@

```{r fig2b}
# Matrix of plots showcasing inter-regional trade trends
fig2b <- artis_region_method %>% 
  ggplot(aes(x = year, y = live_weight_t/1000000, fill = habitat_method)) +
  geom_area() +
  scale_fill_manual(values = habitat_source_colors) +
  labs(x = "", y = "Trade (million t, live weight)", color = "Ave. annual change in export \n(1000 t live weight)") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1), 
        legend.position = "none",
        strip.background = element_blank(),
        plot.margin=unit(c(0,0,0,0), "cm")) +
  facet_grid(rows = vars(exporter_region), cols = vars(importer_region), 
             labeller = label_wrap_gen(width = 5), scales = "free")

@@

```{r fig 2c}
# Figure 2c
# Exports by Region
fig2c <- artis_region %>% 
  ggplot(aes(x = year, y = live_weight_t/1000000, fill = importer_region)) +
  geom_area() +
  labs(x = "", y = "", fill = "Importer") +
  scale_fill_manual(values = region_colors) +
  facet_wrap(~exporter_region, ncol = 1) +
  theme_bw() +
  theme(legend.position = "none",
        strip.background = element_blank(),
        strip.text.x = element_blank(),
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
        plot.margin=unit(c(0,0.5,0,0), "cm"))

@@

```{r fig2}
# Overall Figure 2
ggarrange(ggarrange(fig2a, NULL,
                    ncol = 2, nrow = 1,
                    widths = c(4.5, 1.25), labels = c("a", "")),
          ggarrange(fig2b, 
                    ggarrange(NULL, fig2c,
                              ncol = 1, heights = c(0.3, 4)), 
                    ncol = 2, nrow = 1,
                    widths = c(4.5, 1), labels = c("b", "c")),
          ncol = 1, nrow = 2, 
          heights = c(1, 4.5))

@@

fig3b <- fig3b_data %>%
  ggplot(aes(x = year, y = supply_per_cap, fill = habitat_method)) +
  geom_area() + 
  labs(y = "Per capita supply (kg)", x = "", fill = "Source") +
  scale_fill_manual(values = habitat_source_colors) +
  theme_bw() +
  theme(legend.position = "bottom",
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  facet_wrap(~region)

@@

fig3d <- fig3d_data %>%
  ggplot(aes(x = year, y = supply_percent, color = supply_source)) +
  geom_line() +
  scale_color_manual(values = c("#F37542", "#0F2D59")) +
  labs(x = "", y = "Percent of supply", color = "Origin") +
  facet_wrap(~region) +
  theme_bw() +
  theme(legend.position = "bottom",
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

@@

```{r fig4a}
fig4a_data <- read.csv(file.path(datadir, "figa4a_data.csv"))

@@

top_exporters_old <- read.csv(file.path(datadir, "top_exporters_old.csv"), row.names = FALSE)
top_exporters_recent <- read.csv(file.path(datadir, "top_exporters_recent.csv"), row.names = FALSE)

@@

top_importers_old <- read.csv(file.path(datadir, "top_importers_old.csv"), row.names = FALSE)
top_importers_recent <- read.csv(file.path(datadir, "top_importers_recent.csv"), row.names = FALSE)

@@

```{r si_export_decreases_increases, fig.height = 4, fig.width = 4}
top_exporter_increases <- read.csv(file.path(datadir, "top_export_increases.csv"))
top_exporter_decreases <- read.csv(file.path(datadir, "top_export_decreases.csv"))

@@

```{r si_import_decreases_increases, fig.height = 4, fig.width = 4}
top_importer_increases <- read.csv(file.path(datadir, "top_import_increases.csv"))
top_importer_decreases <- read.csv(file.path(datadir, "top_import_decreases.csv"))

@@

>>>>

```{r load_data}
# Set directories
datadir <- "outputs"
outdir <- "outputs"
```

@@

@@

fig1b <- ggplot() +
  geom_line(data = fig1b_data, aes(x = year, y = percent_export, color = habitat_method)) +
  scale_color_manual(values = habitat_source_colors) +
  labs(y = "% Production Exported", x = "", color = "Source") +
  theme_bw()

@@

```{r fig2a_v2}

region_net_import <- artis_region_method %>% 
  group_by(year, importer_region) %>%
  summarise(import_live_weight_t = sum(live_weight_t, na.rm = TRUE)) %>%
  rename(region = importer_region) %>%
  left_join(artis_region_method %>% 
              group_by(year, exporter_region) %>%
              summarise(export_live_weight_t = sum(live_weight_t, na.rm = TRUE)) %>%
              rename(region = exporter_region), by = c("year", "region")) %>%
  mutate(net_import = import_live_weight_t-export_live_weight_t)

fig2a <- ggplot() +
  geom_area(data = artis_region_method %>% 
              select(-exporter_importer) %>%
              pivot_longer(exporter_region:importer_region, 
                           names_to = c("flow", "drop"), 
                           values_to = "region", names_sep = "_") %>% 
              filter(flow == "exporter") %>%
              group_by(year, region, habitat_method) %>%
              summarise(live_weight_t = sum(live_weight_t, na.rm = TRUE)), 
            aes(x = year, y = -1*live_weight_t/1000000, fill = habitat_method)) +
  geom_area(data = artis_region_method %>% 
              select(-exporter_importer) %>%
              pivot_longer(exporter_region:importer_region, 
                           names_to = c("flow", "drop"), 
                           values_to = "region", names_sep = "_") %>% 
              filter(flow == "importer") %>%
              group_by(year, region, habitat_method) %>%
              summarise(live_weight_t = sum(live_weight_t, na.rm = TRUE)), 
            aes(x = year, y = live_weight_t/1000000, fill = habitat_method)) +
  scale_fill_manual(values = habitat_source_colors) +
  geom_hline(yintercept = 0, color = "black", linewidth = 0.5, linetype = "dashed") +
  geom_line(data = region_net_import, aes(x = year, y = net_import/1000000, color = "Net import"), linewidth = 1) +
  scale_color_manual(values = "black") +
  labs(x = "", y = "Trade (million t, live weight)", fill = "Source", color = "") +
  guides(fill = guide_legend(nrow = 2)) +
  theme_bw() +
  theme(legend.position = "top",
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
        strip.background = element_blank()) +
  facet_wrap(~region, nrow = 1)

@@

```{r fig2a_old, eval=FALSE}
# Imports by Region
fig2a <- artis_region %>% 
  ggplot(aes(x = year, y = live_weight_t/1000000, fill = exporter_region)) +
  geom_area() +
  labs(x = "", y = "", fill = "Exporter") +
  scale_fill_manual(values = region_colors) +
  facet_wrap(~importer_region, nrow = 1) +
  theme_bw() +
  theme(legend.position = "none",
        strip.background = element_blank(),
        strip.text.x = element_blank(),
        axis.text.x = element_blank(),
        plot.margin=unit(c(0.5,0,0,0), "cm"))

```{r fig2b}
# Matrix of plots showcasing inter-regional trade trends
fig2b <- artis_region_method %>% 
  ggplot(aes(x = year, y = live_weight_t/1000000, fill = habitat_method)) +
  geom_area() +
  scale_fill_manual(values = habitat_source_colors) +
  labs(x = "", y = "Trade (million t, live weight)", color = "Ave. annual change in export \n(1000 t live weight)",
       fill = "Source") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1), 
        legend.position = "none",
        strip.background = element_blank()) +
  facet_grid(rows = vars(exporter_region), cols = vars(importer_region), 
             labeller = label_wrap_gen(width = 5), scales = "free")

@@

```{r fig2c, eval=FALSE}
# Figure 2c
# Exports by Region
fig2c <- artis_region %>% 
  ggplot(aes(x = year, y = live_weight_t/1000000, fill = importer_region)) +
  geom_area() +
  labs(x = "", y = "", fill = "Importer") +
  scale_fill_manual(values = region_colors) +
  facet_wrap(~exporter_region, ncol = 1) +
  theme_bw() +
  theme(legend.position = "none",
        strip.background = element_blank(),
        strip.text.x = element_blank(),
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
        plot.margin=unit(c(0,0.5,0,0), "cm"))

@@

```{r fig2_old, eval=FALSE}
# Overall Figure 2
ggarrange(ggarrange(fig2a, NULL,
                    ncol = 2, nrow = 1,
                    widths = c(4.5, 1.25), labels = c("a", "")),
          ggarrange(fig2b, 
                    ggarrange(NULL, fig2c,
                              ncol = 1, heights = c(0.3, 4)), 
                    ncol = 2, nrow = 1,
                    widths = c(4.5, 1), labels = c("b", "c")),
          ncol = 1, nrow = 2, 
          heights = c(1, 4.5))

```{r fig2, fig.height=5, fig.width=4}
# Overall Figure 2
ggarrange(ggarrange(fig2a, NULL,
                    ncol = 2, nrow = 1,
                    widths = c(4.5, 0.2), labels = c("a", "")), 
          fig2b, labels = c("", "b"),
          ncol = 1, heights = c(1.75, 4))

@@

fig3b <- fig3b_data %>%
  ggplot(aes(x = year, y = supply_per_cap, fill = habitat_method)) +
  geom_area() + 
  labs(y = "Per capita supply (kg)", x = "", fill = "Source") +
  scale_fill_manual(values = habitat_source_colors) +
  theme_bw() +
  theme(legend.position = "bottom",
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
        strip.background = element_blank()) +
  facet_wrap(~region)

@@

fig3d <- fig3d_data %>%
  ggplot(aes(x = year, y = supply_percent, color = supply_source)) +
  geom_line() +
  scale_color_manual(values = c("#F37542", "#0F2D59")) +
  labs(x = "", y = "Percent of supply", color = "Origin") +
  facet_wrap(~region) +
  theme_bw() +
  theme(legend.position = "bottom",
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
        strip.background = element_blank())

@@

```{r fig4a}
fig4a_data <- read.csv(file.path(datadir, "fig4a_data.csv"))

@@

top_exporters_old <- read.csv(file.path(datadir, "top_exporters_old.csv"))
top_exporters_recent <- read.csv(file.path(datadir, "top_exporters_recent.csv"))

@@

top_importers_old <- read.csv(file.path(datadir, "top_importers_old.csv"))
top_importers_recent <- read.csv(file.path(datadir, "top_importers_recent.csv"))

@@

```{r si_export_decreases_increases, fig.height = 4, fig.width = 4}
top_export_increases <- read.csv(file.path(datadir, "top_export_increases.csv"))
top_export_decreases <- read.csv(file.path(datadir, "top_export_decreases.csv"))

@@

```{r si_import_decreases_increases, fig.height = 4, fig.width = 4}
top_import_increases <- read.csv(file.path(datadir, "top_import_increases.csv"))
top_import_decreases <- read.csv(file.path(datadir, "top_import_decreases.csv"))

@@

```{r ldc_net_exports_habitat_method}
ldc_net_habitat_method_summary <- read.csv(file.path(datadir, "ldc_net_habitat_method_summary.csv"))

ldc_net_exports_by_source_plot <- ldc_net_habitat_method_summary %>%
  ggplot(aes(x = year, y = net_export/1000000, color = habitat_method)) +
  geom_line() +
  scale_color_manual(values = habitat_source_colors) +
  labs(x = "", y = "Net export (million t)", color = "") +
  theme_minimal() +
  theme(legend.position = "bottom")

ldc_net_exports_by_source_plot
```

```{r trade_concentration_n_countries}
# Concentration of trade by production source, measured as the number of countries comprising 75%
cumulative_percent_threshold <- 75

export_n_countries_concentration <- read.csv(file.path(datadir, "export_concentration_n_countries.csv"))
import_n_countries_concentration <- read.csv(file.path(datadir, "import_concentration_n_countries.csv"))

export_n_countries_concentration_by_source <- read.csv(file.path(datadir, "export_concentration_n_countries_by_source.csv"))
import_n_countries_concentration_by_source <- read.csv(file.path(datadir, "import_concentration_n_countries_by_source.csv"))

export_n_countries_by_source_plot <- ggplot() +
  geom_line(data = export_n_countries_concentration_by_source, aes(x = year, y = n, color = habitat_method)) +
  scale_color_manual(values = habitat_source_colors) +
  geom_line(data = export_n_countries_concentration, aes(x = year, y = n), color = "black") +
  labs(x = "", y = paste("Number of countries comprising ", cumulative_percent_threshold, "% of exports", sep = ""), color = "Source") +
  lims(y = c(0, 30)) +
  theme_bw()

export_n_countries_by_source_plot

import_n_countries_by_source_plot <- ggplot() +
  geom_line(data = import_n_countries_concentration_by_source, aes(x = year, y = n, color = habitat_method)) +
  scale_color_manual(values = habitat_source_colors) +
  geom_line(data = import_n_countries_concentration, aes(x = year, y = n), color = "black") +
  labs(x = "", y = paste("Number of countries comprising ", cumulative_percent_threshold, "% of imports", sep = ""), color = "Source") +
  lims(y = c(0, 30)) +
  theme_bw()

import_n_countries_by_source_plot
```

### 2023-02-04 ms_results.Rmd
Update analyses to exclude non-human food

```{r trends_summary}
# Generate stats for text
total_1996 <- round(sum(global_export_by_source %>% filter(year == 1996) %>% pull(live_weight_t))/1000000, 1)
total_2019 <- round(sum(global_export_by_source %>% filter(year == 2019) %>% pull(live_weight_t))/1000000, 1)

@@

# Total trade for all aquaculture
total_aquaculture_1996 <- round(sum(global_export_by_source %>% filter(year == 1996, str_detect(habitat_method, "aquaculture")) %>% pull(live_weight_t))/1000000, 1)
total_aquaculture_2019 <- round(sum(global_export_by_source %>% filter(year == 2019, str_detect(habitat_method, "aquaculture")) %>% pull(live_weight_t))/1000000, 1)

# Total trade for FMFO
total_fmfo_1996 <- round(sum(global_export_by_source %>% filter(year == 1996, food_or_fmfo == "fishmeal") %>% pull(live_weight_t))/1000000, 1)
total_fmfo_2019 <- round(sum(global_export_by_source %>% filter(year == 2019, food_or_fmfo == "fishmeal") %>% pull(live_weight_t))/1000000, 1)
```

@@

Globalization describes the degree of international connectedness, which can be characterized by the increasing flow of goods and number of trade relationships. Blue foods became increasingly globalized from 1996-2019, with exports increasing `r round(100*((total_2019-total_1996)/total_1996),0)`%  (from `r total_1996` to `r total_2019` mil t) and the average number of export partners increasing `r round(100*((all_degree_2019-all_degree_1996)/all_degree_1996),0)`%  (from `r round(all_degree_1996, 1)` to `r round(all_degree_2019, 1)`) since 1996. Over that period, both farmed and wild exports increased, though aquaculture exports grew faster, more than tripling, whereas capture exports grew by `r round(100*((total_capture_2019-total_capture_1996)/total_capture_1996),0)`%. However, despite aquaculture now comprising half of blue food production, capture fishery products still dominate blue food exports, with capture products responsible for `r 100*round(total_capture_2019/total_2019, 1)`% of exports. Marine-sourced products dominate both capture and aquaculture exports, comprising `r round(100*total_marine_capture_2019/(total_marine_capture_2019+total_inland_capture_2019), 1)`% and `r round(100*total_marine_aquaculture_2019/(total_marine_aquaculture_2019+total_inland_aquaculture_2019), 1)`%, respectively. From 1996 to 2019, fishmeal exports remained relatively stable, declining `r abs(round(100*(total_fmfo_2019-total_fmfo_1996)/(total_fmfo_1996), 1))`% over the time period. Consequently, the share of global blue food exports represented by fishmeal declined from `r abs(round(100*(total_fmfo_1996)/(total_1996), 1))`% in 1996 to only `r abs(round(100*(total_fmfo_2019)/(total_2019), 1))`% in 2019.

Another measure of the degree of globalization is the share of production exported. Domestic exports increased to `r round(percent_export_total_dom_2019, 1)`% of production in 2019 while total exports (i.e., domestic plus exports of foreign and unknown origin) reached `r round(percent_export_total_2019, 1)`% of all production. For comparison, the share of cereal production exported grew from around 10% in the late 1990s to 17% in the 2020s (Clapp 2022). Marine capture products have the greatest share of production destined for export (domestic exports represented `r round(percent_marine_capture_exported_2019, 1)`% of production) and experienced the largest increases in the share exported, while inland aquaculture had the lowest share of production destined for export (domestic exports represented only `r round(percent_inland_aqua_exported_2019, 1)`% of production) (Fig 1b).

One concern for increasing globalization is that it may create systemic risks from shocks, particularly when trade is highly concentrated. Since 1996, blue food exports and imports have become somewhat less concentrated, with only `r export_concentration_n_countries %>% filter(year == 1996) %>% pull(n)` countries comprising `r cumulative_percent_threshold`% of exports in 1996 versus `r export_concentration_n_countries %>% filter(year == 2019) %>% pull(n)` countries in 2019 and `r import_concentration_n_countries %>% filter(year == 1996) %>% pull(n)` countries comprising `r cumulative_percent_threshold`% of imports in 1996 versus `r import_concentration_n_countries %>% filter(year == 2019) %>% pull(n)` countries in 2019. Compare this with crops where just 7 countries and the EU account for 90% of wheat exports and just four countries accounting >80% of maize exports (Clapp 2022). Aquaculture exports are generally concentrated among fewer countries than capture exports, corresponding to the high concentration of aquaculture production. However, inland aquaculture imports have become less concentrated over time. The more dispersed trade patterns can be attributed to expanding middle classes and urbanization, particularly in low- and middle-income countries, which are responsible for increasing blue food demand in many countries (Naylor et al. 2021; SOFIA 2022). [Roz - maybe we can come up with an illustrative example from the BFA demand paper for here?]

![Figure 1: Increases in global export of blue foods. a) Exports of global marine and freshwater aquaculture and fishery products (t live weight equivalent) from 1996-2019. b) Percent of global marine and freshwater aquaculture and fishery production exported, excluding re-exports. c) Out degree by production method and environment.](outputs/fig1.png)

@@

@@

The top importers, exporters, and bilateral flows vary by production method, underscoring the importance of disaggregating blue food trade (Fig SX). In recent years, the marine capture and aquaculture trade networks have been dominated by Asia and Europe, and to a lesser extent, North America (Fig SX). Since 1996, the marine trade networks shifted more toward Asia and Europe, with large increases in marine capture imports and exports by Africa and large increases in marine aquaculture exports from South America and imports to North America (Fig SX). At the country level, although some countries rank among the top traders across production methods, such as China for exports and the USA for imports, many countries are only top traders for one production method. Meanwhile, inland farmed and wild production is oriented more toward domestic consumption and what is exported tends to stay within the region, particularly within Asia (Fig 2; Fig SX). Inland blue food trade is dominated by aquaculture, with the largest exports from Vietnam and China and the largest imports by the United States, Japan, and South Korea (Fig SX).

Intraregional trade is generally stronger due to shorter transport distances, historical ties, and established regional trade agreements (REF). This pattern generally holds for blue food trade as intraregional trade is either the highest or second highest for all regions other than Oceania and South America. [Suggestions for discussion points around Oceania and South America exports?] Regional trade strengthened within Asia, Europe and Africa, where intraregional trade increased more than interegional trade.

Since 1996, nearly all regional trade pairs increased, other than South America to Europe and Asia and within North America. The largest average annual growth increases occurred for trade within Asia, within Europe and between Europe and Asia. At the country level, `r round(percent_flow_increases, 1)`% of trade pairs increased, with the most increases for marine/inland aquaculture/capture. The declining exports from South America mirror declining marine catch over this period. [Suggestions for discussion points and context related to this?].

Least developed countries are net importers of blue foods [true of all production methods?] and this pattern has increased X% since 1996 (Fig SX - need to create this figure). Net imports of blue foods has previously been identified as a threat to nutrition security in these nations as [check ref for reasoning here] (REF). High net imports is also associated with risk of receiving shocks propagated through trade (Gephart et al. 2016).

>>>>

```{r trends_summary}
# Generate stats for text
total_1996 <- round(sum(global_export_by_source %>% filter(year == 1996) %>% pull(live_weight_t))/1000000, 1)
total_2019 <- round(sum(global_export_by_source %>% filter(year == 2019) %>% pull(live_weight_t))/1000000, 1)
total_2020 <- round(sum(global_export_by_source %>% filter(year == 2020) %>% pull(live_weight_t))/1000000, 1)

@@

# Total trade for all capture
total_capture_1996 <- round(sum(global_export_by_source %>% filter(year == 1996, str_detect(habitat_method, "capture")) %>% pull(live_weight_t))/1000000, 1)
total_capture_2019 <- round(sum(global_export_by_source %>% filter(year == 2019, str_detect(habitat_method, "capture")) %>% pull(live_weight_t))/1000000, 1)
total_capture_2020 <- round(sum(global_export_by_source %>% filter(year == 2020, str_detect(habitat_method, "capture")) %>% pull(live_weight_t))/1000000, 1)

# Total trade for all aquaculture
total_aquaculture_1996 <- round(sum(global_export_by_source %>% filter(year == 1996, str_detect(habitat_method, "aquaculture")) %>% pull(live_weight_t))/1000000, 1)
total_aquaculture_2019 <- round(sum(global_export_by_source %>% filter(year == 2019, str_detect(habitat_method, "aquaculture")) %>% pull(live_weight_t))/1000000, 1)
total_aquaculture_2020 <- round(sum(global_export_by_source %>% filter(year == 2020, str_detect(habitat_method, "aquaculture")) %>% pull(live_weight_t))/1000000, 1)

# Total trade for FMFO
artis_fm <- read.csv(file.path(datadir, "artis_fm.csv"))

total_fmfo_1996 <- artis_fm %>%
  filter(year == 1996) %>%
  mutate(live_weight_t = live_weight_t/1000000) %>%
  pull(live_weight_t)

total_fmfo_2019 <- artis_fm %>%
  filter(year == 2019) %>%
  mutate(live_weight_t = live_weight_t/1000000) %>%
  pull(live_weight_t)

@@

Globalization describes the degree of international connectedness, which can be characterized by the increasing flow of goods and number of trade relationships. Aquatic foods became increasingly globalized from 1996-2019, with exports increasing `r round(100*((total_2019-total_1996)/total_1996),0)`%  (from `r total_1996` to `r total_2019` mil t; Fig 1a) and the average number of export partners increasing `r round(100*((all_degree_2019-all_degree_1996)/all_degree_1996),0)`%  (from `r round(all_degree_1996, 1)` to `r round(all_degree_2019, 1)`; Fig 4c) since 1996. Over that period, both farmed and wild exports increased, though aquaculture exports grew faster, more than tripling, whereas capture exports grew by `r round(100*((total_capture_2019-total_capture_1996)/total_capture_1996),0)`%. Corresponding with the first year of the COVID-19 pandemic, global aquatic food exports declined `r -1*round(100*((total_2020-total_2019)/total_2019),1)`% overall in 2020 relative to 2019, with a `r -1*round(100*((total_capture_2020-total_capture_2019)/total_capture_2019),1)`% decline in capture exports, but a `r round(100*((total_aquaculture_2020-total_aquaculture_2019)/total_aquaculture_2019),1)`% increase in aquaculture exports (Fig 1a). Despite aquaculture now comprising half of blue food production, capture fishery products still dominate blue food exports, with capture products responsible for `r 100*round(total_capture_2019/total_2019, 1)`% of exports. Marine-sourced products dominate both capture and aquaculture exports, comprising `r round(100*total_marine_capture_2019/(total_marine_capture_2019+total_inland_capture_2019), 1)`% and `r round(100*total_marine_aquaculture_2019/(total_marine_aquaculture_2019+total_inland_aquaculture_2019), 1)`%, respectively. Aquatic food trade outpaced fishmeal (for non-human consumption) trade, which remained relatively stable, declining `r abs(round(100*(total_fmfo_2019-total_fmfo_1996)/(total_fmfo_1996), 1))`% over the time period. Consequently, fishmeal exports as a share of the sum of aquatic food and fishmeal exports declined from `r abs(round(100*(total_fmfo_1996)/(total_1996+total_fmfo_1996), 1))`% in 1996 to only `r abs(round(100*(total_fmfo_2019)/(total_2019+total_fmfo_2019), 1))`% in 2019.

Another measure of the degree of globalization is the share of production exported. Domestic exports, which represent exports sourced from that nation’s production irrespective of where their production occurred, increased to `r round(percent_export_total_dom_2019, 1)`% of production in 2019 while total exports, which represents domestic plus exports of foreign and unknown origin, reached `r round(percent_export_total_2019, 1)`% of all production. For comparison, the share of cereal production exported grew from around 10% in the late 1990s to 17% in the 2020s (Clapp 2022). Marine capture products have the greatest share of production destined for export (domestic exports represented `r round(percent_marine_capture_exported_2019, 1)`% of production) and experienced the largest increases in the share exported, while inland aquaculture had the lowest share of production destined for export (domestic exports represented only `r round(percent_inland_aqua_exported_2019, 1)`% of production) (Fig 1b).

Globalization may create systemic risks from shocks, particularly when trade is highly concentrated. Since 1996, aquatic food exports and imports have become moderately less concentrated, with only `r export_concentration_n_countries %>% filter(year == 1996) %>% pull(n)` countries comprising `r cumulative_percent_threshold`% of exports in 1996 versus `r export_concentration_n_countries %>% filter(year == 2019) %>% pull(n)` countries in 2019. Compare this with crops where just 7 countries and the EU account for 90% of wheat exports and just four countries accounting >80% of maize exports (Clapp 2022). The declining concentration is driven by capture fishery exports, whereas both inland and marine aquaculture exports became somewhat more concentrated (Fig SX). Aquaculture exports are generally concentrated among fewer countries than capture exports, corresponding to the high concentration of aquaculture production. Import concentration declined more dramatically, with `r import_concentration_n_countries %>% filter(year == 1996) %>% pull(n)` countries comprising `r cumulative_percent_threshold`% of imports in 1996 versus `r import_concentration_n_countries %>% filter(year == 2019) %>% pull(n)` countries in 2019 (Fig SX). The more dispersed import patterns are likely associated with expanding middle classes and urbanization, particularly in low- and middle-income countries, which are responsible for increasing aquatic food demand in many countries (Naylor et al. 2021; SOFIA 2022).

![Figure 1: Increases in global export of aquatic foods. a) Exports of global marine and freshwater aquaculture and fishery products (t live weight equivalent) from 1996-2019. b) Percent of global marine and freshwater aquaculture and fishery production exported, excluding re-exports. c) Out degree by production method and environment.](outputs/fig1.png)

@@

Aquatic foods have become increasingly globalized, with the percent of production exported increasing by `r 100*round((percent_export_total_dom_2019-percent_export_total_dom_1996)/percent_export_total_dom_1996, 1)` and the volume and number of trade partnerships approximately doubling. However, trends differ across aquatic food groups, with marine capture remaining the most highly globalized group, but with marine aquaculture showing faster trade growth.

@@

The top importers, exporters, and bilateral flows vary by production method, underscoring the importance of disaggregating aquatic food trade (Fig SX). In recent years, the marine capture and aquaculture trade networks have been dominated by Asia and Europe, and to a lesser extent, North America (Fig SX). Since 1996, the marine trade networks shifted more toward Asia and Europe, with large increases in marine capture imports and exports by Africa and large increases in marine aquaculture exports from South America and imports to North America (Fig SX). At the country level, although some countries rank among the top traders across production methods, such as China for exports and the USA for imports, many countries are only top traders for one production method. Meanwhile, inland farmed and wild production is oriented more toward domestic consumption and what is exported tends to stay within the region, particularly within Asia (Fig 2; Fig SX). Inland aquatic food trade is dominated by aquaculture, with the highest exports from Vietnam and China and highest imports by the United States, Japan, and South Korea (Fig SX).

Intraregional trade tends to be higher than interregional trade due to shorter transport distances, historical ties, and established regional trade agreements (REF). This pattern generally holds for aquatic food trade as intraregional trade is the highest or second highest for all regions other than Oceania and South America. Our disaggregated trade network is based on a reconciled version of reported trade (Gaulier and Zignago 2010) and therefore does not capture informal and unreported trade networks that often exist among neighboring countries, notably in Africa (Benassi et al. 2019). Since 1996, trade increased between nearly all regional trade pairs, other than South America to Europe and Asia and within North America. At the country level, trade increased between `r round(percent_flow_increases, 1)`% of trade pairs. Despite trade increasing with partners across the globe, regional trade within Asia, Europe and Africa outpaced interregional trade with those regions. The largest average annual growth increases occurred for trade within Asia and within Europe, followed by trade between Europe and Asia.

Least developed countries, as defined by the UN, are net exporters of aquatic foods across all production methods, with net exports more than tripling between 1996 and their peak in 2018 (Fig SX). The pattern is largely driven by marine capture products, which nearly quadrupled, while other aquatic food sector net exports remained relatively stable. Net exports of aquatic foods may be beneficial to least developed countries where high value species are exported and revenue used to purchase other foods (Asche et al, 2015). However, recent work exploring movement of nutrients derived from fisheries suggests international trade is driving redistribution of essential micronutrients prevalent and bioavailable in aquatic foods from areas of high deficiency in middle and low income countries to developed nations with greater nutrient security (Nash et al, 2022). Further, Nash et al (2022) found evidence of high vulnerability to future shifts in trade practices among middle and low income nations currently benefiting from trade in aquatic foods; high net imports are associated with the risk of receiving shocks propagated through trade (Gephart et al. 2016).
